While a thesis is ostensibly the work of a single author, it is the worst guarded secret of academia that no thesis would be possible without the tireless support of many, many people.
Even though the work itself is mine, it would not have been possible without advisors, mentors, co-authors, friends, and family.

Of course, the first and biggest ``thank you'' is addressed to my supervisors who have guided me along my path for the last five years.
Without all of you, I would not be here!
Amir-massoud Farahmand, thank you for pushing me constantly and tirelessly to become more precises and clear in my thinking and writing.
Animesh Garg, thank you for your enthusiasm and strong vision, and the many amazing small and large tips and tricks you have passed on to me to structure my writing and presentation.
Igor Gilitschenski, thank you for stepping up when I needed help and being the person I could always turn to with my frustrations and emotions.
Additional thanks of course goes to the amazing chair of my supervisory committee and mentor.
Wil Cunningham, thank you for your raw enthusiasm, curiosity and playfulness in exploration.
One day, I believe you will truly know how the brain works!

Getting to a PhD program is in itself a long road, and so I want to thank the professors I had the great pleasure of learning from during my undergrad and master's.
A special thanks goes of course to my Darmstadt thesis supervisor Kristian Kersting, and the two professors who mentored me and who trusted my abilities enough to write the reference letters that got me to Toronto, Jan Peters, and Johannes Fuernkranz.

The next round of thanks goes out to the collaborators and friends I have in Toronto and beyond.
To my coauthors Romina Abachi, Tyler Kastner, Victor Liao, Arash Ahmadian, Anastasiia Pedan, Jonas Guan, Eric Eaton, and last but definitely not least, Marcel Hussing, thanks for all of your amazing work!
Maria and Taylor, I promise you you will be on this list soon enough, so I am including you in spirit already.
To the undergraduate students who worked with me, Victor Liao, Arash Ahmadian, Yash Ramani, Allan Lin, Jasper Gerig, Anastasiia Pedan, Tehilla Helfenbaum, Ian Vyse, and Shabnam Tajik, thanks for putting up with my meandering ideas and good luck on all of your paths.
Finally, even though we have also not yet officially co-authored a paper, a big thanks Eugene Vinitsky for becoming a great collaborator and friend over the final stretch of my PhD.

A big thanks goes to Jonathan Tompson for hosting me at DeepMind, and to Heni Ben-Amor for wonderful coffee chats.
And of course, a special thank you to Maria Attarian for being the best office mate!

The next group of people is the biggest part of the acknowledgements, my dear friends and lab mates.
Roughly in order of appearance, I thank
the Adage lab and visitors, Amin, Romina, Avery, Mete, Tyler and Sierra, Amarildo, and Jongmin; the (late) PAIR lab, Maria and Jon, Kate, Lily, Ziyi, Mayank, and Dylan; the TISL lab and visitors, Ashkan and Sarah, Sam, Yash, Umangi, Evgenii, Jasper, Mohammad, Kai, Justin, Alfred, Shuhong, Andrei, Toshi, Axel, Eugen, and Marcel; the CleverHans lab or extended Nicola mafia or potluck gang and visitors, Stefan, Mohammad, Avital, Sierra, Anvith, Patty, and David; friends from around Vector, Lukas, Taylor, Sep, Shalev, and Andreas; Abdur Mohammed and Jackie Zuborg representative for the Vector staff who were always quick to help with any problems; and finally, the great friends from all around the University of Toronto, especially Silvia and Gabriel, Varun, and Arthur.

Some friends have remained close after I moved to Toronto, and so a special thanks goes to Marcel, the D120 and WiMi gang, Stephan and Maria, Julian and Gesine, Nadja, Benjamin, Jannis, Anna and Philip, Fabian and Steffi, Johny and Franzi, and Sial, Robyn, Kai and family.

Collectively, I want to thank Queer in AI for their incredible work, and I send out all my love to those who find it harder to fight for the right to live their life in truth every day.
Love and community will prevail in the end.

My love goes out to one of my most influential teachers, Dorothee, requiescat in pace.
I had hoped so much that I would be able to show you this.
You pushed me since high school to accomplish great things and both your wise words and cutting mockery were always perfectly balanced to keep me on track.

A special thanks goes to Corinne, for helping to keep me sane in Toronto, especially during the dark days of lock-downs, remote learning and teaching, and a long-distance relationship.

Of course, there are not enough words to express my thanks to my family, especially my parents, Nikola and Hinrich, and my brother and sister-in-law, Till and Jasmine.
Without your love, encouragement, and support I would have never managed to get this far (both in terms of work and physical distance).

Finally, for Heiko, I have no words to adequately express everything I owe you.
Instead of words, you get me for the rest of your life.
